Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-11-28 analysis/paper/paper.qmd
Edited conclusion

Considering the importance of stemmed points for the technological transition during the Late Paleolithic in Korea, our research examines the likely uses of stemmed points by asking three research questions: What were stemmed points used for? How diverse were their likely uses? What are the temporal and spatial patterns in stemmed point use? We applied tip cross-sectional area (TCSA) metric because it has been used as a ballistically relevant standard to discriminate different likely use classes of projectile points and it requires only a few measurements on a stone artifact (i.e width and thickness) to compute the metric. We calculated TCSA for a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites. Then we examined the TCSA values with other variables including raw materials, weight, radiocarbon dates, and site locations. Based on the evolutionary approach, we premised that the way of using stemmed points is likely to reflect the socio-environmental circumstances that people encountered and managed. We examined the possible impact of LGM on stemmed points and their distribution in eco-regional maps.

According to the different weapon-delivery systems that summarize TCSA of projectile points from archaeological and ethnographic cases, a majority of stemmed points might be used the most likely as javelin tips and stabbing spear tips. But overall we observed a wide range and also difference distribution of TCSA in each assemblage. Therefore, we conclude that stemmed points might play diverse roles during the Late Paleolithic. Based on our results, we assume that people were exposed to largely unfamiliar situation and created a stemmed point as multifunctional tool to performed multiple tasks. We are aware that discriminating the likely use of small projectile points could be arbitrary [@erlandson2014darts]. Since TCSA covers the critical elements of projectiles, flight and penetration dynamics (i.e. increased or decrease by shape of tip and cross section), we still consider it a useful metric for discriminating between weapon tip tools [@hughes1998getting; @sitton2020tip; @lombard2021variation]. In our future research, we will conduct usewear analysis on stemmed points and experimental research as another attempt to identify detailed function of stemmed points and the technological connection between hunter-gatherer groups.

>>>>

Considering the importance of stemmed points for the technological transition during the Late Paleolithic in Korea, our research examines the likely uses of stemmed points by asking three research questions: What were stemmed points used for? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We applied the tip cross-sectional area (TCSA) metric because it has been used as a ballistically relevant standard to discriminate different likely use classes of projectile points and it requires only a few measurements on a stone artifact (i.e. width and thickness) to compute the metric. We calculated TCSA for a total of `r min_n_points` stemmed points from `r n_ass` assemblages unearthed from `r n_sites` sites. Then we examined the TCSA values with other variables including raw materials, weight, radiocarbon dates, and site locations. Based on the evolutionary approach, we premised that the way of using stemmed points is likely to reflect the socio-environmental circumstances that people encountered and managed. We examined the possible impact of LGM on stemmed points and their distribution on eco-regional maps.

According to the different weapon-delivery systems that summarize TCSA of projectile points from archaeological and ethnographic cases, a majority of stemmed points might be used the most likely as javelin tips and stabbing spear tips. In general, though, we noted a wide range and also differing distribution of TCSA in each assemblage. Therefore, we conclude that stemmed points might play diverse roles during the Late Paleolithic. From our findings, we infer that people were exposed to an unfamiliar situation and used a stemmed point as a multifunctional tool to carry out multiple tasks. We are aware that discriminating the likely use of small projectile points could be arbitrary [@erlandson2014darts]. Since TCSA covers the critical elements of projectiles, flight and penetration dynamics (i.e. increased or decrease by shape of tip and cross section), we still consider it a useful metric for discriminating between weapon tip tools [@hughes1998getting; @sitton2020tip; @lombard2021variation]. As part of our future research, we will conduct usewear analyses of stemmed points and experimental research in an effort to elucidate the detailed function of stemmed points and technological links between hunter-gatherer groups.
